### 2020-05-06 writeup/transfer.tex
edits

\section{Variance Reduction from Information Transfer}\label{sec_denoise}

Recall that the test error of $\hat{\beta}_{t}^{\MTL}$ consists of two parts
\begin{align}
	\te(\hat{\beta}_t^{\MTL}) = \hat{w}^2 \bignorm{\Sigma_2^{1/2} (\hat{w}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}X_1^{\top}X_1 (\beta_s - \hat{w}\beta_t)}^2 + \sigma^2\cdot \bigtr{(\hat{w}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}\Sigma_2}. \label{eq_te_model_shift}
\end{align}
It is not hard to show that the variance of $\hat{\beta}_t^{\MTL}$ is reduced compared to $\hat{\beta}_t^{\STL}$ (following the argument of Proposition \ref{prop_monotone}), i.e.
\[ \sigma^2\cdot \bigtr{(\hat{w}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}\Sigma_2} \le \sigma^2\cdot \bigtr{(X_2^{\top}X_2)^{-1}\Sigma_2}. \]
Because of model shift however, i.e. $\beta_s \neq \beta_t$.
We can no longer guarantee that $\te(\hat{\beta}_{t}^{\MTL}) \le \te(\hat{\beta}_t^{\STL})$.
The main result of this part show deterministic conditions under which we get positive or negative transfer.
And the conditions depend only on the covariate shift matrix $M$, the difference of the task models, and the number of per-task data points.
In order to characterize $\te(\hat{\beta}_t^{\MTL})$ and $\te(\hat{\beta}_t^{\STL})$, the technical crux of our approach relies on deriving the limit of the trace of matrix inverse in the high-dimensional setting.
To illustrate the idea, we observe that by using Lemma \ref{lem_minv}, we have that
\[ \te(\hat{\beta}_t^{\STL}) = \frac{\sigma^2}{n_2 - p}\bigtr{\Sigma_2^{-1}}. \]
We shall also derive the limit of $\te(\hat{\beta}_t^{\MTL})$.
\todo{write a brief technical overview}

\begin{algorithm}[!t]
	\caption{Comparing the test error of MTL, TL and STL estimators}
	\begin{algorithmic}[1]
		\Input Two regression tasks $(X_1, Y_2)$, $(X_2, Y_2)$.
		\Param Shared body $B$, task-specific prediction heads $W_1, W_2$.
		\State Training the shared body $B$.
		\State Optimizing the task heads on the validation set

		\begin{itemize}
			\item Jointly optimizing both tasks: $\hat{\beta}_t^{\MTL}$
			\item Optimizing the target task: $\hat{\beta}_t^{\TL}$
			\item Single-task training baseline: $\hat{\beta}_t^{\STL}$
		\end{itemize}
		\State Problem statement: Provide conditions to compare the test error of the three estimators on the target task, $\te(\hat{\beta}_t^{\MTL})$, $\te(\hat{\beta}_t^{\STL})$ and $\te(\hat{\beta}_t^{\TL})$.
	\end{algorithmic}
\end{algorithm}

\begin{theorem}\label{thm_model_shift}
	Let $n_1, n_2$ be the number of data points for the source, target task, respectively.
	Let $\hat{w}$ denote the optimal solution for the ratio $w_1/w_2$ in equation \eqref{eq_val_mtl}.
	Let ${M} = \hat{w} \Sigma_1^{1/2}\Sigma_2^{-1/2}$ denote the weighted covariate shift matrix.
	Denote by ${\lambda}_1, {\lambda}_2, \dots, {\lambda}_p$ the singular values of ${M}^{\top}{M}$.
	The information transfer is solely determined by two deterministic quantities $\Delta_{\beta}$ and $\Delta_{\vari}$, which show the change of model shift bias and variace, respectively.
	With high probability we have
	\begin{align}
	 	\te(\hat{\beta}_{t}^{\MTL}) \le \te(\hat{\beta}_t^{\STL}) \text{ when: } &\Delta_{\vari} - \Delta_{\beta} \ge \left(\bigbrace{1 + \sqrt{\frac{p} {n_1}}}^4 - 1 \right) \delta \label{upper}\\
		\te(\hat{\beta}_t^{\MTL}) \ge \te(\hat{\beta}_t^{\STL}) \text{ when: } &\Delta_{\vari} - \Delta_{\beta} \le -2\bigbrace{2\sqrt{\frac {p}{n_1}} + \frac{p}{n_1}} \delta, \label{lower}
	\end{align}
	where $\delta = \norm{Z} \cdot \norm{\Sigma_1^{1/2}(\beta_s - \hat{w}\beta_t)}^2$ with $Z = \frac{n_1^2}{(n_1 + n_2)^2}\cdot{M} \frac{(1 + a_3)\id + a_4 {M}^{\top}{M}}{(a_2 + a_1 {M}^{\top}{M})^2} {M}^{\top}$, and
	\begin{align*} %\bigtr{{\Sigma_2^{-1}}}
		\Delta_{\vari} &\define {\sigma^2}\bigbrace{\frac{p}{n_2 - p} -  \frac{1}{n_1 + n_2} \bigtr{(a_1 M^{\top}M + a_2\id)^{-1}} } \\
		\Delta_{\beta} &\define (\beta_s - \hat{w}\beta_t)^{\top} \Sigma_1^{1/2} Z \Sigma_1^{1/2} (\beta_s - \hat{w}\beta_t),
	\end{align*}
%	{\color{blue}To write the bounds in the form \eqref{upper} and \eqref{lower}, we have to define $\wh w$ as the minimizer of
%	$$\frac{\sigma^2}{n_1 + n_2} \bigtr{\frac1{a_1(w) M(w)^{\top}M(w) + a_2(w)\id}}+ \Delta_\beta(w).$$
%	Otherwise, the two bounds has to be written into the form
%	\begin{align*}
%	{\sigma^2}\frac{p}{n_2 - p}\ge \min_w \left\{\frac{\sigma^2}{n_1 + n_2} \bigtr{\frac1{a_1(w) M(w)^{\top}M(w) + a_2(w)\id}} + \Delta_\beta(w)+ \left(\bigbrace{1 + \sqrt{\frac{p} {n_1}}}^4 - 1 \right) \delta(w)\right\},
%	\end{align*}
%	and
%	\begin{align*}
%	{\sigma^2}\frac{p}{n_2 - p}\le \min_w \left\{\frac{\sigma^2}{n_1 + n_2} \bigtr{\frac1{a_1(w) M(w)^{\top}M(w) + a_2(w)\id}}+ \Delta_\beta(w) - 2\bigbrace{2\sqrt{\frac {p}{n_1}} + \frac{p}{n_1}} \delta(w)\right\}.
%	\end{align*}
%	}
%
%
	and $a_1, a_2, a_3, a_4$ are the solutions of the following linear equations
	\begin{gather}
		 a_1 + a_2 = 1- \frac{p}{n_1 + n_2},~ a_1 + \sum_{i=1}^p \frac{a_1}{(n_1 + n_2)(a_1 + a_2/ \lambda_i^2)} = \frac{n_1}{n_1 + n_2}, \label{eq_a2} \\
		\left(\frac{n_2}{a_2^2}- \sum_{i=1}^p \frac{1}{ (a_2 + \lambda_i^2a_1)^2  }\right) a_3 -  \left(\sum_{i=1}^p \frac{  \lambda_i^2 }{ (  a_2 + \lambda_i^2a_1)^2  }\right)a_4
		= \sum_{i=1}^p \frac{1 }{ (  a_2 + \lambda_i^2a_1)^2  }, \label{eq_a3} \\
		\left(\frac{n_1}{a_1^2} -  \sum_{i=1}^p \frac{\lambda_i^4   }{  (a_2 + \lambda_i^2a_1)^2  }\right)a_4 -\left(\sum_{i=1}^p \frac{\lambda_i^2  }{  (a_2 + \lambda_i^2a_1)^2  }\right)a_3
		= \sum_{i=1}^p \frac{\lambda_i^2 }{  (a_2 + \lambda_i^2a_1)^2  }. \label{eq_a4}
	\end{gather}
\end{theorem}

%\todo{For technical reasons, we will consider the test error.}
%\section{Effects of Model Distance, Covariate Shift, and Data Ratio on Information Transfer}\label{sec_insight}
\subsection{Effect of Model Distance on Positive versus Negative Transfer}

@@

\section{Effect of Covariate Shift and Data Ratio on Transfer Rate}

@@

@@

>>>>

\section{Illustrative Examples}

\subsection{Effect of Model Distance on Positive versus Negative Transfer}

@@

\subsection{Effect of Covariate Shift and Data Ratio on Transfer Rate}

@@

Note that for the case of $k$ tasks with the same covariates, since there is no covariate shift and the data ratio is always equal to one, the main factor is model distance.

@@

\section{Main Results: Reducing Variance under Model Shift Biases}\label{sec_denoise}

Recall that the test error of $\hat{\beta}_{t}^{\MTL}$ consists of two parts
\begin{align}
	\te(\hat{\beta}_t^{\MTL}) = \hat{w}^2 \bignorm{\Sigma_2^{1/2} (\hat{w}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}X_1^{\top}X_1 (\beta_s - \hat{w}\beta_t)}^2 + \sigma^2\cdot \bigtr{(\hat{w}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}\Sigma_2}. \label{eq_te_model_shift}
\end{align}
It is not hard to show that the variance of $\hat{\beta}_t^{\MTL}$ is reduced compared to $\hat{\beta}_t^{\STL}$ (following the argument of Proposition \ref{prop_monotone}), i.e.
\[ \sigma^2\cdot \bigtr{(\hat{w}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}\Sigma_2} \le \sigma^2\cdot \bigtr{(X_2^{\top}X_2)^{-1}\Sigma_2}. \]
Because of model shift however, i.e. $\beta_s \neq \beta_t$.
We can no longer guarantee that $\te(\hat{\beta}_{t}^{\MTL}) \le \te(\hat{\beta}_t^{\STL})$.
The main result of this part show deterministic conditions under which we get positive or negative transfer.
And the conditions depend only on the covariate shift matrix $M$, the difference of the task models, and the number of per-task data points.
In order to characterize $\te(\hat{\beta}_t^{\MTL})$ and $\te(\hat{\beta}_t^{\STL})$, the technical crux of our approach relies on deriving the limit of the trace of matrix inverse in the high-dimensional setting.
To illustrate the idea, we observe that by using Lemma \ref{lem_minv}, we have that
\[ \te(\hat{\beta}_t^{\STL}) = \frac{\sigma^2}{n_2 - p}\bigtr{\Sigma_2^{-1}}. \]
We shall also derive the limit of $\te(\hat{\beta}_t^{\MTL})$.
\todo{write a brief technical overview}
We introduce these quantities, $Z = \frac{n_1^2}{(n_1 + n_2)^2}\cdot{M} \frac{(1 + a_3)\id + a_4 {M}^{\top}{M}}{(a_2 + a_1 {M}^{\top}{M})^2} {M}^{\top}$, and
	\begin{align*} %\bigtr{{\Sigma_2^{-1}}}
		\Delta_{\vari} &\define {\sigma^2}\bigbrace{\frac{p}{n_2 - p} -  \frac{1}{n_1 + n_2} \bigtr{(a_1 M^{\top}M + a_2\id)^{-1}} } \\
		\Delta_{\beta} &\define (\beta_s - \hat{w}\beta_t)^{\top} \Sigma_1^{1/2} Z \Sigma_1^{1/2} (\beta_s - \hat{w}\beta_t),
	\end{align*}
%	{\color{blue}To write the bounds in the form \eqref{upper} and \eqref{lower}, we have to define $\wh w$ as the minimizer of
%	$$\frac{\sigma^2}{n_1 + n_2} \bigtr{\frac1{a_1(w) M(w)^{\top}M(w) + a_2(w)\id}}+ \Delta_\beta(w).$$
%	Otherwise, the two bounds has to be written into the form
%	\begin{align*}
%	{\sigma^2}\frac{p}{n_2 - p}\ge \min_w \left\{\frac{\sigma^2}{n_1 + n_2} \bigtr{\frac1{a_1(w) M(w)^{\top}M(w) + a_2(w)\id}} + \Delta_\beta(w)+ \left(\bigbrace{1 + \sqrt{\frac{p} {n_1}}}^4 - 1 \right) \delta(w)\right\},
%	\end{align*}
%	and
%	\begin{align*}
%	{\sigma^2}\frac{p}{n_2 - p}\le \min_w \left\{\frac{\sigma^2}{n_1 + n_2} \bigtr{\frac1{a_1(w) M(w)^{\top}M(w) + a_2(w)\id}}+ \Delta_\beta(w) - 2\bigbrace{2\sqrt{\frac {p}{n_1}} + \frac{p}{n_1}} \delta(w)\right\}.
%	\end{align*}
%	}
%
%
	and $a_1, a_2, a_3, a_4$ are the solutions of the following linear equations
	\begin{gather}
		 a_1 + a_2 = 1- \frac{p}{n_1 + n_2},~ a_1 + \sum_{i=1}^p \frac{a_1}{(n_1 + n_2)(a_1 + a_2/ \lambda_i^2)} = \frac{n_1}{n_1 + n_2}, \label{eq_a2} \\
		\left(\frac{n_2}{a_2^2}- \sum_{i=1}^p \frac{1}{ (a_2 + \lambda_i^2a_1)^2  }\right) a_3 -  \left(\sum_{i=1}^p \frac{  \lambda_i^2 }{ (  a_2 + \lambda_i^2a_1)^2  }\right)a_4
		= \sum_{i=1}^p \frac{1 }{ (  a_2 + \lambda_i^2a_1)^2  }, \label{eq_a3} \\
		\left(\frac{n_1}{a_1^2} -  \sum_{i=1}^p \frac{\lambda_i^4   }{  (a_2 + \lambda_i^2a_1)^2  }\right)a_4 -\left(\sum_{i=1}^p \frac{\lambda_i^2  }{  (a_2 + \lambda_i^2a_1)^2  }\right)a_3
		= \sum_{i=1}^p \frac{\lambda_i^2 }{  (a_2 + \lambda_i^2a_1)^2  }. \label{eq_a4}
	\end{gather}

\begin{theorem}\label{thm_model_shift}
	Let $n_1, n_2$ be the number of data points for the source, target task, respectively.
	Let $\hat{w}$ denote the optimal solution for the ratio $w_1/w_2$ in equation \eqref{eq_val_mtl}.
	Let ${M} = \hat{w} \Sigma_1^{1/2}\Sigma_2^{-1/2}$ denote the weighted covariate shift matrix.
	Denote by ${\lambda}_1, {\lambda}_2, \dots, {\lambda}_p$ the singular values of ${M}^{\top}{M}$.
	The information transfer is solely determined by two deterministic quantities $\Delta_{\beta}$ and $\Delta_{\vari}$, which show the change of model shift bias and variace, respectively.
	Let $\delta = \norm{Z} \cdot \norm{\Sigma_1^{1/2}(\beta_s - \hat{w}\beta_t)}^2$.
	With high probability we have
	\begin{align}
	 	\te(\hat{\beta}_{t}^{\MTL}) \le \te(\hat{\beta}_t^{\STL}) \text{ when: } &\Delta_{\vari} - \Delta_{\beta} \ge \left(\bigbrace{1 + \sqrt{\frac{p} {n_1}}}^4 - 1 \right) \delta \label{upper}\\
		\te(\hat{\beta}_t^{\MTL}) \ge \te(\hat{\beta}_t^{\STL}) \text{ when: } &\Delta_{\vari} - \Delta_{\beta} \le -2\bigbrace{2\sqrt{\frac {p}{n_1}} + \frac{p}{n_1}} \delta, \label{lower}
	\end{align}
\end{theorem}

%\todo{For technical reasons, we will consider the test error.}
%\section{Effects of Model Distance, Covariate Shift, and Data Ratio on Information Transfer}\label{sec_insight}
